---
title: "Logic Models"
subtitle: "Proof of Concept for Creating a Logic Model in graphviz"
author: "Andy Grogan-Kaylor" 
date: "today"
format:
  html:
    toc: true
    number-sections: true
    code-fold: true
    code-summary: "Show the code"
    theme:
      light: cosmo
      dark: vapor
---

# Introduction

Logic models are used extensively in program development and program evaluation.

This is an evolving proof of concept of creating a logic model in graphviz. [graphviz](https://graphviz.org/) is a programming language designed to represent graphs. graphviz files end in `.gv`.

graphviz is implemented in [RStudio](https://www.rstudio.com/) via the [DiagrammeR](https://rich-iannone.github.io/DiagrammeR/) package. This logic model was made with RStudio.

This proof of concept makes use of *subgraphs* to organize the components of a logic model.

Like all programming languages, graphviz may facilitate automation, replication, and documentation of a project. In this way, using graphviz may make creating a logic model easier.

Like all programming languages, graphviz may be finicky and hard to tweak. In this way, using graphviz may make creating a logic model much more difficult. 

**It may be much easier to use PowerPoint or Google Slides to create a logic model!!!**

# Call the `DiagrammeR` Library

```{r}

library(DiagrammeR)

```

# Render the Graph of the Logic Model

The graphviz code for the logic model can be found [here](https://github.com/agrogan1/teaching/blob/master/logic-models/logic-model.gv). 

```{r}

DiagrammeR::grViz("logic-model.gv")

```

# You Can Use RStudio to Work Directly in graphviz.

As long as you have installed `DiagrammeR` (`install.packages("DiagrammeR")`) you can work directly with a `.gv` file without working in R. You are welcome to download this [sample `logic-model.gv` file](https://github.com/agrogan1/teaching/blob/master/logic-models/logic-model.gv) to experiment with this.







