---
title: "Logic Models"
description: |
  Proof of Concept for Creating a Logic Model in graphviz.
author:
  - name: Andy Grogan-Kaylor 
    url: https://agrogan1.github.io/
    affiliation: University of Michigan
    affiliation_url: https://umich.edu/
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    highlight: haddock
    toc: true
    toc_float: true
    code_folding: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

# Introduction

Logic models are used extensively in program development and program evaluation.

This is an evolving proof of concept of creating a logic model in graphviz. [graphviz](https://graphviz.org/) is a programming language designed to represent graphs. graphviz files end in `.gv`.

graphviz is implemented in [RStudio](https://www.rstudio.com/) via the [DiagrammeR](https://rich-iannone.github.io/DiagrammeR/) package. This logic model was made with RStudio.

This proof of concept makes use of *subgraphs* to organize the components of a logic model.

Like all programming languages, graphviz may facilitate automation, replication, and documentation of a project. In this way, using graphviz may make creating a logic model easier.

Like all programming languages, graphviz may be finicky and hard to tweak. In this way, using graphviz may make creating a logic model much more difficult. 

**It may be much easier to use PowerPoint or Google Slides to create a logic model!!!**

# Call the `DiagrammeR` Library

```{r}

library(DiagrammeR)

```

# Render the Graph of the Logic Model

The graphviz code for the logic model can be found [here](logic-model.gv). 

```{r}

DiagrammeR::grViz("logic-model.gv")

```





