---
title: "Logic Models"
description: |
  Proof of Concept for Creating a Logic Model in graphviz.
author:
  - name: Andy Grogan-Kaylor 
    url: https://agrogan1.github.io/
    affiliation: University of Michigan
    affiliation_url: https://umich.edu/
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    highlight: haddock
    toc: true
    toc_float: false
    code_folding: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

Logic models are used extensively in program development and program evaluation.

This is an evolving proof of concept of creating a logic model in graphviz. [graphviz](https://graphviz.org/) is a programming language designed to represent graphs. graphviz files end in `.gv`.

graphviz is implemented in [RStudio](https://www.rstudio.com/) via the [DiagrammeR](https://rich-iannone.github.io/DiagrammeR/) package. This logic model was made with RStudio.

This proof of concept makes use of *subgraphs* to organize the components of a logic model.

Like all programming languages, graphviz may facilitate automation, replication, and documentation of a project. In this way, using graphviz may make creating a logic model easier.

Like all programming languages, graphviz may be finicky and hard to tweak. In this way, using graphviz may make creating a logic model much more difficult. 

**It may be much easier to use PowerPoint or Google Slides to create a logic model!!!**

```{r}

library(DiagrammeR)

```

```{r}

grViz("

digraph logicmodel {
  
  graph [rankdir = LR]
  
  label = 'Logic Model'

  subgraph cluster_0 {
  	    
  	label = 'Major Domains'
    
    workshops [label='Workshops']

    publications [label='Publications']

  	}
  	
  	subgraph cluster_1 {
  	    
  	label = 'Activities'
  	
  	meetings [label = 'Meetings']
  	
  	workshops -> meetings
    
  	}
  	
  	subgraph cluster_2 {
  	    
  	label = 'Outcomes'
  	
  	better_parenting [label = 'Better Parenting']
  	
  	meetings -> better_parenting
  	
  	publications -> better_parenting
    
  	}

}

")
```





