---
title: "Quantitative Data Analysis"
description: |
  In Excel, Google Sheets, R, and Stata
author: Andy Grogan-Kaylor
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    highlight: haddock
    toc: yes
    code_folding: true
  powerpoint_presentation: 
    toc: yes
    fig_caption: yes
  slidy_presentation: 
    fig_caption: yes
    highlight: haddock
    css: UMslidy.css
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

library(pander)

xaringanExtra::use_panelset()

```

# Publicly Available Tools for Analysis

-------------------------------------------------------------
Tool        Cost         Ease        Analysis       Suitability 
                         of          Capabilities   for
                         Use                        Large Data
----------- -------      ----------  -------------  ------------
Excel       Comes        Easy        Limited        Difficult when
            installed                               N > 100
            on many
            computers

Google      Free         Easy        Limited        Difficult when
Sheets      with a                                  N > 100
            Google
            account
            
R           Free         Challenging Extensive      Excellent with
                                                    large datasets
                                                    
Stata       Some         Learning    Extensive      Excellent with 
            cost         Curve                      large datasets
                         (Intuitive) 
-------------------------------------------------------------

# Our Data

We take a look at our *simulated* data, which has an `id` number, `age`, and `happiness` (on a 5 point scale, with 5 being the happiest.)

```{r, echo=FALSE}

N <- 100 # sample size

id <- seq(1, N) # id number

age <- rnorm(N, 50, 10) # normally distributed age

age[2] <- 200 # someone is 200 years old

happy <- round(runif(N, 1, 5)) # 5 point happiness scale

happy[1] <- -99

happy[3] <- -99

somethingelese <- rnorm(N, 0, 1) # something else!

mydata <- data.frame(id, age, happy, somethingelese)

pander(head(mydata))

```
Notice that…

* There are variables in which we may not have interest (e.g. `somethingelse`).
* None of the variables have informative *variable labels*. We have to guess at what the variables mean.
* Variables do not seem to have informative *value labels*. While somewhat intuitive, we have to guess at what the values mean.
* Someone appears to 200 years old.
* There appear to be missing values in the variable happy that need to be recoded.

# Cleaning Data

* Only keep the variables of interest.
* Add variable labels (if we can).
* Add value labels (if we can).
* Recode outliers, values that are errors, or values that should be coded as missing

::::: {.panelset}


::: {.panel}

## Excel {.panel-name}

Testing

:::


::: {.panel}

## Google Sheets {.panel-name}

Testing 1,2,3

:::

:::::

# Simple Analysis

