---
title: "Differential Equations"
description: |
  Reviewing some basic ideas in differential equations.
author:
  - name: Andy Grogan-Kaylor 
    url: https://www.umich.edu/~agrogan
    affiliation: University of Michigan
    affiliation_url: https://www.umich.edu
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    highlight: haddock
    toc: yes
    number_sections: yes
    fig_height: 3
  tint::tintPdf:
    highlight: haddock
    number_sections: yes
    toc: yes
  tufte::tufte_handout:
    toc: yes
    number_sections: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      fig.margin = TRUE)

library(ggplot2)

library(gganimate)

library(xaringan)

library(xaringanExtra)

xaringanExtra::use_panelset()


```


# Background

In this handout, we consider several simple differential equations that characterize growth. We then consider the solution of this differential equation.

# Constant (Linear) Growth

```{r}

x <- seq(1,10)

y1 <- as.numeric(x)

mydata1 <- data.frame(x, y1)

plinear <- ggplot(data = mydata1,
                  aes(x = x,
                      y = y1,
                      group = 1)) +
  # geom_point() +
  geom_line(size = 2) +
  labs(title = "Constant Growth") +
  transition_reveal(x)

```

::::: {.panelset}

::: {.panel}

## $dN/dt$ Form {.panel-name}

$$\frac{dN}{dt} = k$$ 

implies 

$$N = kt$$

```{r}

plinear + labs(subtitle = "Population Size Grows Linearly As A Function Of Time",
               x = "t", 
               y = "N")

```

:::

::: {.panel}

## $dy/dx$ Form {.panel-name}

$$\frac{dy}{dx} = k$$ 

implies 

$$y = kx + C$$

```{r}

plinear + labs(subtitle = "y Grows Linearly As A Function Of x")

```

:::

:::::

# Exponential Growth

```{r}

x <- seq(1,10)

y2 <- exp(x)

mydata2 <- data.frame(x, y2)

pexponential <- ggplot(data = mydata2,
                       aes(x = x,
                           y = y2,
                           group = 1)) +
  # geom_point() +
  geom_line(size = 2) +
  labs(title = "Exponential Growth") +
  transition_reveal(x)


```

::::: {.panelset}

::: {.panel}

## $dN/dt$ Form {.panel-name}

$$\frac{dN}{dt} = y$$

implies

$$N = e^t$$

```{r}

pexponential  + labs(subtitle = "Population Size Grows Exponentially As A Function Of Time",
                     x = "t", 
                     y = "N")

```

:::

::: {.panel}

## $dy/dx$ Form {.panel-name}

$$\frac{dy}{dx} = y$$

implies

$$y = e^x$$

```{r}

pexponential  + labs(subtitle = "y Grows Exponentially As A Function Of x")

```

:::

:::::


# Logistic Growth

## $dy/dx$ Form

## $dN/dt$ Form





