---
title: "Differential Equations"
description: |
  Reviewing some basic ideas in differential equations.
author:
  - name: Andy Grogan-Kaylor 
    url: https://www.umich.edu/~agrogan
    affiliation: University of Michigan
    affiliation_url: https://www.umich.edu
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    highlight: haddock
    toc: yes
    number_sections: yes
    fig_height: 4
  tint::tintPdf:
    highlight: haddock
    number_sections: yes
    toc: yes
  tufte::tufte_handout:
    toc: yes
    number_sections: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      fig.margin = TRUE)

library(ggplot2)

library(gganimate)

library(plotly)

library(xaringan)

library(xaringanExtra)

xaringanExtra::use_panelset()


```


# Background

In this handout, we consider several simple differential equations that characterize growth. We then consider the solution of each differential equation in terms of $y$ as a function of $x$.

I consider each differential equation in 2 different--mathematically equivalent--forms: 

* A $dN/dt$ form where it is sometimes easier to consider the *substantive implications* by thinking about population growth over time.
* A $dy/dx$ form where it is sometimes easier to think about the *solution* to the differential equation.

# Constant (Linear) Growth

There is a constant rate of growth, expressed by $k$.

```{r}

x <- seq(1,10)

y1 <- as.numeric(x)

mydata1 <- data.frame(x, y1)

plinear <- ggplot(data = mydata1,
                  aes(x = x,
                      y = y1,
                      group = 1)) +
  geom_line(size = 2, color = "grey") +
  geom_point(aes(frame = x), color = "deepskyblue", size = 2) +
  labs(title = "Constant Growth") # +
  # transition_reveal(x)

```

::::: {.panelset}

::: {.panel}

## $dN/dt$ Form {.panel-name}

$$\frac{dN}{dt} = k$$ 

implies 

$$N = kt + C$$

```{r}

plinearA <- plinear + 
  labs(subtitle = "Population Size Grows Linearly As A Function Of Time",
       x = "t", 
       y = "N")

ggplotly(plinearA)

```

:::

::: {.panel}

## $dy/dx$ Form {.panel-name}

$$\frac{dy}{dx} = k$$ 

implies 

$$y = kx + C$$

```{r}

plinearB <- plinear + 
  labs(subtitle = "y Grows Linearly As A Function Of x")

ggplotly(plinearB)

```

:::

:::::

# Exponential Growth

Growth is a function of population size $N$ (or equivalently $y$).

```{r}

x <- seq(1, 10, .5)

y2 <- exp(x)

mydata2 <- data.frame(x, y2)

pexponential <- ggplot(data = mydata2,
                       aes(x = x,
                           y = y2,
                           group = 1)) +
  geom_line(size = 2, color = "grey") +
  geom_point(aes(frame = x), color = "deepskyblue", size = 2) +
  labs(title = "Exponential Growth") # +
  # transition_reveal(x)


```

::::: {.panelset}

::: {.panel}

## $dN/dt$ Form {.panel-name}

$$\frac{dN}{dt} = N$$

implies

$$N = e^t + C$$

```{r}

pexponentialA <- pexponential + 
  labs(subtitle = "Population Grows Exponentially As A Function Of Time",
       x = "t", 
       y = "N")

ggplotly(pexponentialA)

```

:::

::: {.panel}

## $dy/dx$ Form {.panel-name}

$$\frac{dy}{dx} = y$$

implies

$$y = e^x + C$$

```{r}

pexponentialB <- pexponential + 
  labs(subtitle = "y Grows Exponentially As A Function Of x",
       y = "y")

ggplotly(pexponentialB)

```

:::

:::::

# Logistic Growth

Growth is initially approximately exponential, but then tapers off as $N$ or $y$ approaches some limiting value.

```{r}

x <- seq(1, 20)

y3 <- exp(x - 10) / (1 + exp(x - 10))

mydata3 <- data.frame(x, y3)

plogistic <- ggplot(data = mydata3,
                       aes(x = x,
                           y = y3,
                           group = 1)) +
  geom_line(size = 2, color = "grey") +
  geom_point(aes(frame = x), color = "deepskyblue", size = 2) +
  labs(title = "Logistic Growth") # +
  # transition_reveal(x)


```


::::: {.panelset}

::: {.panel}

## $dN/dt$ Form {.panel-name}

$$\frac{dN}{dt} = N \left(1 - \frac{N}{N_0} \right)$$

implies

$$y = N_0 \left( \frac{e^t}{1 + e^t} \right)$$

```{r}

plogisticA <- plogistic + 
  labs(subtitle = "Population Grows Exponentially Until It Reaches Carrying Capacity",
       x = "t", 
       y = "N") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

ggplotly(plogisticA)

```

:::

::: {.panel}

## $dy/dx$ Form {.panel-name}

$$\frac{dy}{dx} = y \left(1 - y \right)$$
implies

$$y = \frac{e^x}{1 + e^x}$$

```{r}

plogisticB <- plogistic + 
  labs(subtitle = "y Grows Exponentially Until y Reaches 1.0",
       y = "y")

ggplotly(plogisticB)

```

:::

:::::


