---
title: "Differential Equations"
subtitle: "Reviewing some basic ideas in differential equations."
author: "Andy Grogan-Kaylor"
date: "today"
format: 
  html:
    toc: true
    number-sections: true
    anchor-sections: true
    theme:
      light: cosmo
      dark: vapor
    code-fold: true
    code-summary: "Show the code"
---

```{r setup, include=FALSE}

library(ggplot2)

library(gganimate)

library(plotly)

```

# Background

In this handout, we consider several simple differential equations that characterize growth. We then consider the solution of each differential equation in terms of $y$ as a function of $x$.

For *logistic growth*, I also consider a $dp/dx$ formulation.

# Constant (Linear) Growth

There is a constant rate of growth, expressed by $k$.

```{r}
#| warning: false

x <- seq(1,10)

y1 <- as.numeric(x)

mydata1 <- data.frame(x, y1)

plinear <- ggplot(data = mydata1,
                  aes(x = x,
                      y = y1,
                      group = 1)) +
  geom_line(linewidth = 2, color = "grey") +
  geom_point(aes(frame = x), color = "deepskyblue", size = 3) +
  labs(title = "Constant Growth") 

```


$$\frac{dy}{dx} = k$$ 

implies 

$$y = kx + C$$

```{r}

plinearB <- plinear + 
  labs(subtitle = "y Grows Linearly As A Function Of x",
       y = "y")

ggplotly(plinearB)

```

# Exponential Growth

Growth is a function of $y$.

```{r}
#| warning: false

x <- seq(1, 10, .5)

y2 <- exp(x)

mydata2 <- data.frame(x, y2)

pexponential <- ggplot(data = mydata2,
                       aes(x = x,
                           y = y2,
                           group = 1)) +
  geom_line(linewidth = 2, color = "grey") +
  geom_point(aes(frame = x), color = "deepskyblue", size = 3) +
  labs(title = "Exponential Growth") 

```

$$\frac{dy}{dx} = y$$

implies

$$y = e^x + C$$

```{r}

pexponentialB <- pexponential + 
  labs(subtitle = "y Grows Exponentially As A Function Of x",
       y = "y")

ggplotly(pexponentialB)

```

# Logistic Growth

Growth is initially approximately exponential, but then tapers off as $y$ approaches some limiting value.

```{r}
#| warning: false

x <- seq(1, 20)

y3 <- exp(x - 10) / (1 + exp(x - 10))

mydata3 <- data.frame(x, y3)

plogistic <- ggplot(data = mydata3,
                       aes(x = x,
                           y = y3,
                           group = 1)) +
  geom_line(linewidth = 2, color = "grey") +
  geom_point(aes(frame = x), color = "deepskyblue", size = 3) +
  labs(title = "Logistic Growth") 

```


::: {.panel-tabset}

## $dy/dx$ Form 

$$\frac{dy}{dx} = y \left(1 - y \right)$$
implies

$$y = \frac{e^x}{1 + e^x} + C$$

```{r}

plogisticB <- plogistic + 
  labs(subtitle = "y Grows Exponentially Until y Approaches 1.0",
       y = "y")

ggplotly(plogisticB)

```

## $dp/dx$ Form {.panel-name}

Here $p$ is a probability.

$$\frac{dp}{dx} = p \left(1 - p \right)$$
implies

$$p = \frac{e^x}{1 + e^x} + C$$

```{r}

plogisticC <- plogistic + 
  labs(subtitle = "p Grows Exponentially Until p Approaches 1.0",
       y = "probability")

ggplotly(plogisticC)

```

:::





