---
title: "Mediation"
subtitle: "Trying to understand mediation, and mediation vs. confounding"
author: "Andy Grogan-Kaylor"
date: "today"
format: 
  html:
    theme: 
      light: yeti
      dark: vapor
    toc: true
    number-sections: true
    anchor-sections: true
---

## Mediation

The basic idea is that if we are regressing $y$ on variable of interest $x$, and (not shown) control variable(s) $z$, if we introduce mediator $m$, the $\beta_{xy}$ coefficient should change. Additionally, $\beta_{xm} \times \beta_{my}$ should be statistically significant.[^bootstrap]

[^bootstrap]: The statistical significance of this quantity may need to be calculated with bootstrapping.

```{r}
#| fig-cap: "Mediation"
#| eval: true
#| echo: false

DiagrammeR::grViz("

digraph boxes_and_circles {
  
  # a 'graph' statement
  graph [overlap = true, 
        fontsize = 20,
        rankdir = LR]

  # several 'node' statements
  node [shape = box,
        fontname = Helvetica]

  node [shape = circle,
        fontsize = 20,
        fixedsize = true,
        width = 1.25] // sets as circles

  x1 [label = 'x']
  
  m1 [label = 'm']

  y1 [label = 'y']
  
  # several 'edge' statements
  
 	subgraph cluster_1 {
 	
 	label = 'mediation';
 	
  x1 -> m1 [label = <&beta;<SUB>xm</SUB>>]
  m1 -> y1 [label = <&beta;<SUB>my</SUB>>]
  x1 -> y1 [label = <&beta;<SUB>xy</SUB>>]

 	}  

}
                  
")
```

## Mediation vs. Correlated Predictors

As far as I know, all of the relationships detailed below would result in similar patterns of correlations. As far as I know there would be no empirical way to distinguish between these relationships. Only theory, or knowledge of prior relationships could be used to distinguish between these possible explanations for correlated $x$, $m$ and $y$.

```{r}
#| fig-cap: "Mediation vs. Confounding"
#| eval: true
#| echo: false

DiagrammeR::grViz("

digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = true, 
        fontsize = 20,
        rankdir = LR]

  # several 'node' statements
  node [shape = box,
        fontname = Helvetica]

  node [shape = circle,
        fontsize = 20,
        fixedsize = true,
        width = 1.25] // sets as circles
  
  x1 [label = 'x']
  
  x2 [label = 'x']
  
  m1 [label = 'm']

  m2 [label = 'm']

  y1 [label = 'y']
  
  y2 [label = 'y']
  
  x1 -> x2 [style=invis]

  # several 'edge' statements
 	
 	subgraph cluster_1 {
 	
 	label = 'm mediates the relationship between x and y';

  x1 -> m1 
  m1 -> y1 
  x1 -> y1 
  
 	}  
 	
 	subgraph cluster_2 {
 	  
 	label = 'x and m are correlated predictors of y';
  
  x2 -> m2 [dir=both]
  x2 -> y2
  m2 -> y2
  
  {rank = same; x2 m2 }
  
 	}
  
}


")
```
