---
title: "Regression Models"
description: |
  Possible regression models for dissertation.
author: "Andy Grogan-Kaylor"
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    highlight: haddock
    toc: yes
  pdf_document: 
    toc: yes
    number_sections: yes
    highlight: haddock
    latex_engine: xelatex
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

library(DiagrammeR)

```

# Conceptual Model

```{r}

grViz(  "

digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = true, 
        fontsize = 10,
        rankdir = LR]

  # several 'node' statements
  node [shape = box,
        fontname = Helvetica]

  node [shape = circle,
        fontsize = 11,
        fixedsize = true,
        width = 1.25] // sets as circles
 
  A [label = 'attachment \nbehavior']
  PRF[label = 'parental \nreflective \nfunctioning']
  CR[label = 'caregiving \nrepresentation']
  CB[label = 'caregiving \nbehavior']
  
  PRF -> A[label = 'moderated by \ntreatment']
  CR -> A[label = 'moderated by \ntreatment']
  CB -> A[label = 'moderated by \ntreatment']
  controls -> A
   
}

")

```

# Syntax

Assuming variables `PRF`, `CR`, `CB` and `attachment`. Additionally, `Tx` is an indicator for treatment, and `x`, `y` and `z` represent the control variables.


-------------------------------------------------------------
SPSS[^SPSS]                              Stata[^Stata]
--------------------                     --------------------
`REGRESSION`                             `regress attachment PRF CR CB x y z`

`/MISSING LISTWISE`          

`/STATISTICS COEFF OUTS`

`/CRITERIA=PIN(.05) POUT(.10)`

`/NOORIGIN`

`/DEPENDENT attachment`

`/METHOD=ENTER PRF CR CB Tx x y z`

-------------------------------------------------------------


[^SPSS]: SPSS syntax is constantly in flux, but this is a best guess at appropriate commands.
[^Stata]: [Stata makes things easy](https://agrogan1.github.io/Stata/two-page-stata/TwoPageStata.pdf)! However, if it is not used in your department, there may be too high of a learning cost.




