---
title: "Regression Models"
description: |
  Possible regression models for dissertation.
author: "Andy Grogan-Kaylor"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: yes
    number_sections: yes
    highlight: haddock
    latex_engine: xelatex
  distill::distill_article:
    highlight: haddock
    toc: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

library(DiagrammeR)

```

# Conceptual Model

```{r}

grViz(  "

digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = true, 
        fontsize = 10,
        rankdir = LR]

  # several 'node' statements
  node [shape = box,
        fontname = Helvetica]

  node [shape = circle,
        fontsize = 11,
        fixedsize = true,
        width = 1.25] // sets as circles
 
  A [label = 'attachment \nbehavior']
  PRF[label = 'parental \nreflective \nfunctioning']
  CR[label = 'caregiving \nrepresentation']
  CB[label = 'caregiving \nbehavior']
  
  PRF -> A[label = 'moderated by \ntreatment']
  CR -> A[label = 'moderated by \ntreatment']
  CB -> A[label = 'moderated by \ntreatment']
  controls -> A
   
}

")

```

# Equation

$$\text{attachment} = \beta_0 + \beta_1 \text{PRF} + \beta_2 \text{CR} + \beta_3 \text{CB} + \beta_4 x + \beta_5 z + e_i$$

# Syntax

Assuming variables `PRF`, `CR`, `CB` and `attachment`. Additionally, `Tx` is an indicator for treatment, and `x` and `z` represent the control variables.

## Basic Models

-------------------------------------------------------------
SPSS[^SPSS]                              Stata[^Stata]
--------------------                     --------------------
`REGRESSION`                             `regress attachment PRF CR CB Tx x z`

`/MISSING LISTWISE`                       

`/STATISTICS COEFF OUTS`                  

`/CRITERIA=PIN(.05) POUT(.10)`            

`/NOORIGIN`                                

`/DEPENDENT attachment`                    

`/METHOD=ENTER PRF CR CB Tx x z.`             

-------------------------------------------------------------

## Interactions

If we think there is moderation, we need to include $\beta_6 \text{PRF} * \text{Tx} + \beta_7 \text{CR} * \text{Tx} + \beta_8 \text{CB} * \text{Tx}$ terms in the equation.

-------------------------------------------------------------
SPSS                                     Stata
-----------------------                  --------------------
In SPSS interaction terms                Stata is able to create interactions   
need to be created *by                   *on the fly*.    
hand*.
                                                
`COMPUTE PRF_Tx = PRF * Tx.`             `regress attachment (PRF CR CB)##Tx x z` 
                                          
`COMPUTE CR_Tx = CR * Tx.`

`COMPUTE CB_Tx = CB * Tx.`

These terms would need 
to be entered into the 
equation above.

-------------------------------------------------------------


[^SPSS]: SPSS syntax is constantly in flux, but this is a best guess at appropriate commands. Of course one potential advantage of SPSS is that models can be run using menus.
[^Stata]: [Stata makes things easy](https://agrogan1.github.io/Stata/two-page-stata/TwoPageStata.pdf)! However, if it is not used in your department, there may be too high of a learning cost.




