---
title: "Presenting Interactions"
author: "Andy Grogan-Kaylor"
format: 
  revealjs:
    theme: solarized
    lightbox: true
    smaller: false
    scrollable: true
    html-math-method: mathjax
    transition: slide
    controls: true 
    controls-tutorial: true
    code-fold: true
    code-summary: "show the code"
    chalkboard: true
---

## Background

*x* is an independent variable. *y* is an outcome. *m* is a moderator.

::: {.callout-tip appearance="simple"}
### A Substantive Example

For the sake of illustration, we might imagine that *x* is *bad parenting*, *y* is *mental health problem* or *aggression*, and *m* is *disadvantaged neighborhood* or *disadvantaged context*.
:::

::: {.callout-tip appearance="simple"}
### Dichotomous Variables

This example is presented with dichotomous *x* and *m*, but the same principles would apply with continuous *x* and *m*.
:::

```{r}

library(ggplot2)

library(ggthemes)

```


```{r}

set.seed(1234)

N <- 1000

x <- rbinom(N, 1, .5) # binomial x

m <- rbinom(N, 1, .5) # binomial moderator (m)

e <- rnorm(N, 0, 5) # normal error

y <- x + m + 3 * x * m + e

mydata <- data.frame(x, y, m)

```

## Equation

$$y = \beta_0 + \beta_1 x + \beta_2 m + \beta_3 x \times m + e_i$$

## Independent Variable on Horizontal Axis; Dependent Variable on Vertical Axis; Separate Lines For Moderator

```{r}
#| label: fig-plot1
#| fig-cap: "Independent Variable on Horizontal Axis; Dependent Variable on Vertical Axis; Separate Lines For Moderator"

ggplot(mydata, 
       aes(x = x,
           y = y,
           color = factor(m))) +
  geom_smooth(method = "lm") +
  labs(title = "Outcome by Independent Variable and Moderator") +
  scale_color_solarized(name = "m") +
  theme_solarized() +
  theme(title = element_text(size = rel(1.5)))

```

## Moderator Variable on Horizontal Axis; Dependent Variable on Vertical Axis; Separate Lines For Independent Variable

```{r}
#| label: fig-plot2
#| fig-cap: "Moderator Variable on Horizontal Axis; Dependent Variable on Vertical Axis; Separate Lines For Independent Variable"

ggplot(mydata, 
       aes(x = m,
           y = y,
           color = factor(x))) +
  geom_smooth(method = "lm") +
  labs(title = "Outcome by Independent Variable and Moderator") +
  scale_color_solarized(name = "x") +
  theme_solarized() +
  theme(title = element_text(size = rel(1.5)))

```






